---
title: "illuminance_calculation"
author: "Carolina Guidolin"
date: "2025-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Calculation of illuminance levels for each VR input intensity

Intensities from VR span from 0 to 1, and we are interested in measuring the illuminance received at the eye level at these different input intensities.

### Importing files

```{r}
# Upload necessary packages 
library(tidyverse)

# Setting working directory
setwd(here::here("results")) # specify which folder in the WD the files are to be found 

# Import files (csv) all in one df
raw_data <- list.files(pattern = "\\.csv$") %>%
  purrr::map_df(~read_csv(.))
```

### Data wrangling

When we collected the data (i.e performed the calibration), we only added VR_number after we started the calibration already, meaning that not all files saved contain this variable. We decided to include this variable because we tested 5 VR headsets on only 3 computer, hence using the computer name to extract the VR headset was not sufficient.

```{r}
# Identify missing VR numbers
vr_na <- raw_data %>%
  filter(is.na(VR_number)) # corresponds to PC name RGMS-VR02

# Check that VR_number = 2 does not in fact exist
vr_two <- raw_data %>%
  filter(VR_number == 2)

# Replace missing values 
data1 <- raw_data %>%
  mutate(VR_number = if_else(PC_name == "RGMS-VR02" & is.na(VR_number),
                             2, VR_number
                             )
         )

# Check no NAs left
data1 %>% filter(is.na(VR_number)) # correct, no NA values left 
```

### Now we want to calculate the average illuminance for the left and right eye for each VR headset

We calculate descriptive statistics for each VR headset, for each eye, across the various repetitions of the same input intensity

```{r}
# Removing scientific notation
options(scipen=999)

# Create a for loop to obtain 5 different dfs called vr_[name of vr]
vr_numbers <- c(1,2,3,4,5)

for (vr in vr_numbers) {
  assign(paste0("vr_", vr),
         data1 %>%
         filter(VR_number == vr) %>%
         group_by(eye, input_intensity) %>% 
         summarise(
           mean_photopic = mean(illuminance_lux),
           sd_photopic = sd(illuminance_lux),
           mean_medi = mean(a_opic_iprgc),
           sd_medi = sd(a_opic_iprgc),
           median_photopic = median(illuminance_lux),
           median_medi = median(a_opic_iprgc)
         ) %>%
         ungroup() %>%
         mutate(vr = paste0(vr)) # add extra col to label the vr 
  )
}

```

### Comparison across VR headset

In the last step, we calculated the mean across repetitions for each input intensity. In this step, we compare these values across VR headsets. We are particularly interested in the intensities 0 and 0.5, because these are the intensities for the dark and bright light condition. So we will select these.

```{r}
# Combine dfs into one
vr_combined <- rbind(vr_1, vr_2, vr_3, vr_4, vr_5)

#Df with only dark condition (input intensity = 0)
int0_all <- vr_combined %>%
  filter(input_intensity == 0.0)
  
# Summary stats for the 10 measurements of darkness   
int0_summary <- int0_all %>%
  filter(input_intensity == 0.0) %>%
  summarise(mean = mean(mean_medi),
            sd = sd(sd_medi),
            max = max(mean_medi),
            min = min(mean_medi),
            range = max - min,
            ) %>%
  ungroup()

#Df with bright light condition (input intesity=0.5)
int50_all <- vr_combined %>%
  filter(input_intensity == 0.5)

#Summary stats for the 10 bright light measurements
int50_summary <- int50_all %>%
  summarise(mean = mean(mean_medi),
            sd = sd(sd_medi),
            max = max(mean_medi),
            min = min(mean_medi),
            range = max - min,
            ) %>%
  ungroup()

```

### Turning these into tables for registered report

We now want to create tables to be inserted in the registered report, for both the dark and the bright conditions.

#### Dark condition df processing
```{r}
# First, we need to convert to wide format for nicer visualisation of left/right eye, and combine mean and SD info in one cell 
int0_wide <- int0_all %>%
  select(-c(input_intensity, median_photopic, median_medi)) %>%
  pivot_wider(
    names_from = eye, 
    values_from = c(mean_photopic, sd_photopic, mean_medi, sd_medi)
  ) %>%
  
  # Format cols so that one cell contains mean and sd 
  mutate(photopic_left = paste0(round(mean_photopic_left, 3), "±", round(sd_photopic_left, 3)),
         photopic_right = paste0(round(mean_photopic_right, 3), "±", round(sd_photopic_right, 3)),
         medi_left = paste0(round(mean_medi_left, 3), "±", round(sd_medi_left, 3)),
         medi_right = paste0(round(mean_medi_right, 3), "±", round(sd_medi_right, 3))
  ) %>%
    # Now, we need to calculate the mean of the average and mean of SD for each VR headset (mean for each row). We do this by adding
    # a new col to the df
    mutate(avg_mean_phot_headset = rowMeans(select(., mean_photopic_left, mean_photopic_right),
                                           na.rm = TRUE),
           avg_mean_medi_headset = rowMeans(select(., mean_medi_left, mean_medi_right),
                                           na.rm = TRUE),
           avg_sd_phot_headset = rowMeans(select(., sd_photopic_left, sd_photopic_right),
                                         na.rm = TRUE),
           avg_sd_medi_headset= rowMeans(select(., sd_medi_left, sd_medi_right),
                                         na.rm = TRUE),
           avg_phot_headset = paste0(round(avg_mean_phot_headset, 3), "±", round(avg_sd_phot_headset, 3)),
           avg_medi_headset = paste0(round(avg_mean_medi_headset, 3), "±", round(avg_sd_medi_headset, 3))
  ) 
 # Select only cols that are relevant (combined mean and SD)
int0_wide_clean <- int0_wide %>% 
                   select(vr, photopic_left, photopic_right, medi_left, medi_right, avg_phot_headset, avg_medi_headset) 
  
# Now, we need to calculate the mean of the average and mean of SD across VR headsets, i.e. mean of average and mean of SD for each columns. We do this by adding a final row to the df. This is not so easy as adding a column (using mutate()), so we need to create an additional 1-row df for this
summary_row_0 <- int0_wide %>%
  summarize(
    vr = "Inter-VR averages (mean±1SD)",
    photopic_left = paste0(round(mean(mean_photopic_left, na.rm = TRUE), 3), "±", round(mean(sd_photopic_left, na.rm = TRUE), 3)),
    photopic_right = paste0(round(mean(mean_photopic_right, na.rm = TRUE), 3), "±", round(mean(sd_photopic_right, na.rm = TRUE), 3)),
    medi_left = paste0(round(mean(mean_medi_left, na.rm = TRUE), 3), "±", round(mean(sd_medi_left, na.rm = TRUE), 3)),
    medi_right = paste0(round(mean(mean_medi_right, na.rm = TRUE), 3), "±", round(mean(sd_medi_right, na.rm = TRUE), 3)),
  )

# Append the summary row to the clean dataset
int0_wide_all <- bind_rows(int0_wide_clean, summary_row_0) 

#Replace NA values with empty strings
int0_wide_all <- replace(int0_wide_all, is.na(int0_wide_all), "")


```

Now we create a gt table
```{r}

  
# Use gt package to create a table
table_dark <- int0_wide_all %>%
   gt::gt() %>%
  
  # Merge headers for photopic illuminance and Melanopic EDI
  gt::tab_spanner(
    label = "Photopic illuminance [lux] (mean±1SD)",
    columns = c(photopic_left, photopic_right)
  ) %>%
  gt::tab_spanner(
    label = "Melanopic EDI [lux] (mean±1SD)",
    columns = c(medi_left, medi_right)
  ) %>%
  gt::tab_spanner(
    label= "Intra-VR headset averages (mean±1SD)",
    columns=(c(avg_phot_headset, avg_medi_headset))
  ) %>%
  
  #Center-align the text in each cell
  gt::cols_align(align = "center",
                 columns = dplyr::everything()) %>%
  
  # Rename columns for clarity 
  gt::cols_label(
    vr = "VR",
    photopic_left = "Left eye",
    photopic_right = "Right eye",
    medi_left = "Left eye",
    medi_right = "Right eye",
    avg_phot_headset = "Photopic illuminance",
    avg_medi_headset = "Melanopic EDI"
  ) %>%
  
  # Adding row separator
  gt::tab_style(
    #Row separator format
    style = list(
      cell_borders(
        sides = "bottom",
        color = "grey",
        weight = px(2)
      )
    ),
      locations = list(
        #Location of row separator
        cells_body(
          columns = c(photopic_left, photopic_right, medi_left, medi_right),
          rows = 5)
      )) %>%
  
    # Adding col separator
  gt::tab_style(
    #Col separator format
    style = list(
      cell_borders(
        sides = "right",
        color = "grey",
        weight = px(2)
      )
    ),
      locations = list(
        #Location of col separator
        cells_body(
          columns = c(medi_right))
      )) %>%
  
  # Add title
  gt:: tab_header(title = gt::md("**Stimulus properties for dark condition (<0.01 lx)**"))

table_dark
  
```

#### Bright condition table

We now do the same for the bright condition table.

```{r}
# As before, we first need to convert to wide format for nicer visualisation of left/right eye, and combine mean and SD info in one cell 
int50_wide <- int50_all %>%
  select(-c(input_intensity, median_photopic, median_medi)) %>%
  pivot_wider(
    names_from = eye, 
    values_from = c(mean_photopic, sd_photopic, mean_medi, sd_medi)
  ) %>%
  mutate(photopic_left = paste0(round(mean_photopic_left, 2), "±", round(sd_photopic_left, 2)),
         photopic_right = paste0(round(mean_photopic_right, 2), "±", round(sd_photopic_right, 2)),
         medi_left = paste0(round(mean_medi_left, 2), "±", round(sd_medi_left, 2)),
         medi_right = paste0(round(mean_medi_right, 2), "±", round(sd_medi_right, 2))
  ) %>%
    # Now, we need to calculate the mean of the average and mean of SD for each VR headset (mean for each row). We do this by adding
    # a new col to the df
    mutate(avg_mean_phot_headset = rowMeans(select(., mean_photopic_left, mean_photopic_right),
                                           na.rm = TRUE),
           avg_mean_medi_headset = rowMeans(select(., mean_medi_left, mean_medi_right),
                                           na.rm = TRUE),
           avg_sd_phot_headset = rowMeans(select(., sd_photopic_left, sd_photopic_right),
                                         na.rm = TRUE),
           avg_sd_medi_headset= rowMeans(select(., sd_medi_left, sd_medi_right),
                                         na.rm = TRUE),
           avg_phot_headset = paste0(round(avg_mean_phot_headset, 2), "±", round(avg_sd_phot_headset, 2)),
           avg_medi_headset = paste0(round(avg_mean_medi_headset, 2), "±", round(avg_sd_medi_headset, 2))
  ) 
 # Select only cols that are relevant (combined mean and SD)
int50_wide_clean <- int50_wide %>% 
                   select(vr, photopic_left, photopic_right, medi_left, medi_right, avg_phot_headset, avg_medi_headset) 
  
# Now, we need to calculate the mean of the average and mean of SD across VR headsets, i.e. mean of average and mean of SD for each columns. We do this by adding a final row to the df. This is not so easy as adding a column (using mutate()), so we need to create an additional 1-row df for this
summary_row_50 <- int50_wide %>%
  summarize(
    vr = "Inter-VR averages (mean±1SD)",
    photopic_left = paste0(round(mean(mean_photopic_left, na.rm = TRUE), 2), "±", round(mean(sd_photopic_left, na.rm = TRUE), 2)),
    photopic_right = paste0(round(mean(mean_photopic_right, na.rm = TRUE), 2), "±", round(mean(sd_photopic_right, na.rm = TRUE), 2)),
    medi_left = paste0(round(mean(mean_medi_left, na.rm = TRUE), 2), "±", round(mean(sd_medi_left, na.rm = TRUE), 2)),
    medi_right = paste0(round(mean(mean_medi_right, na.rm = TRUE), 2), "±", round(mean(sd_medi_right, na.rm = TRUE), 2)),
  )

# Append the summary row to the clean dataset
int50_wide_all <- bind_rows(int50_wide_clean, summary_row_50)

#Replace NA values with empty strings
int50_wide_all <- replace(int50_wide_all, is.na(int50_wide_all), "")


```

Create gt table
```{r}
# Use gt package to create a table
table_bright <- int50_wide_all %>%
   gt::gt() %>%
  
  # Merge headers for photopic illuminance and Melanopic EDI
  gt::tab_spanner(
    label = "Photopic illuminance [lux] (mean±1SD)",
    columns = c(photopic_left, photopic_right)
  ) %>%
  gt::tab_spanner(
    label = "Melanopic EDI [lux] (mean±1SD)",
    columns = c(medi_left, medi_right)
  ) %>%
  gt::tab_spanner(
    label= "Intra-VR headset averages (mean±1SD)",
    columns=(c(avg_phot_headset, avg_medi_headset))
  ) %>%
  
  #Center-align the text in each cell
  gt::cols_align(align = "center",
                 columns = dplyr::everything()) %>%
  
  # Rename columns for clarity 
  gt::cols_label(
    vr = "VR",
    photopic_left = "Left eye",
    photopic_right = "Right eye",
    medi_left = "Left eye",
    medi_right = "Right eye",
    avg_phot_headset = "Photopic illuminance",
    avg_medi_headset = "Melanopic EDI"
  ) %>%
  
  # Adding row separator
  gt::tab_style(
    #Row separator format
    style = list(
      cell_borders(
        sides = "bottom",
        color = "grey",
        weight = px(2)
      )
    ),
      locations = list(
        #Location of row separator
        cells_body(
          columns = c(photopic_left, photopic_right, medi_left, medi_right),
          rows = 5)
      )) %>%
  
    # Adding col separator
  gt::tab_style(
    #Col separator format
    style = list(
      cell_borders(
        sides = "right",
        color = "grey",
        weight = px(2)
      )
    ),
      locations = list(
        #Location of col separator
        cells_body(
          columns = c(medi_right))
      )) %>%
  
  # Add title
  gt:: tab_header(title = gt::md("**Stimulus properties for bright condition (~100 lx)**"))

table_bright
  
```

#### Saving both tables

```{r}
# Load chromote, needed to save using gt
library(chromote)

# Make sure you set your path to Chrome using Sys.setenv(CHROMOTE_CHROME = "path/to/chrome.exe") and check this was correct by running chromote::find_chrome()

# Also, make sure to have a folder called "results"!
## The following line ensures that such folder exists
main_dir <-  here::here("results")
if (!dir.exists(main_dir)) dir.create(main_dir, recursive = TRUE)

# Dark condition table
gt::gtsave(table_dark,
           filename = "vr_cali_dark.png",
           path = main_dir,
           vheight = 300, 
           vwidth = 700,
           expand = 15)

# Bright condition table
gt::gtsave(table_bright,
           filename = "vr_cali_bright.png",
           path = main_dir,
           vheight = 300, 
           vwidth = 700,
           expand = 15)
```

#### Alternative visualisation

An alternative to creating two tables is to just create a single table combining both the dark and bright conditions. It's not so easy to stack gt tables, because there is no function in the package to do so. But we borrow a function from [Andrew Weatherman's GitHub repository](https://gist.github.com/andreweatherman/cc17d9b42f922abf8ccaffae0fef0f71).

```{r}
# helper function to stack tables
stack_gt_tables <- function (tables = NULL, output = "viewer", filename = NULL,
                             path = NULL, vwidth = 1000, vheight = 1000, ..., zoom = 2
                          )
{
  # divs for each table // change style from gt_two_column_layout
  stacked_tables <- htmltools::div(
    htmltools::div(tables[[1]], style = "display: block; width: 100%;"),
    htmltools::div(tables[[2]], style = "display: block; width: 100%;")
  )

  # carry rest of gt_two_column_layout
  if (output == "viewer") {
    htmltools::browsable(stacked_tables)
  }
  else if (output == "save") {

    filename <- if (is.null(filename)) tempfile(fileext = ".png") else filename
    filename <- if (is.null(path)) filename else file.path(path, filename)

    tempfile_ <- tempfile(fileext = ".html")
    htmltools::save_html(html = stacked_tables, file = tempfile_)

    webshot2::webshot(url = paste0("file:///", tempfile_),
                      file = filename, vwidth = vwidth, vheight = vheight, zoom = zoom, ...)

  }
  else if (output == "html") {
    stacked_tables
  }
}

# stack_gt_tables(list(table1, table2), output = 'save', file = 'test.png')
```

Now we use this to stack our two tables. Note that from the original function, we slightly changed the vheight and vwidth parameters to fit the current table. We also deleted the zoom and expand parameters
```{r}
dark_bright_combi <- stack_gt_tables(list(table_dark, table_bright),
                                     output = 'save', 
                                     file = 'dark_bright_combined.png',
                                     path = main_dir)
```

