---
title: "jeti_csv_processing"
author: "Carolina Guidolin"
date: "2025-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Processing of csv file obtained from JETI calibration 
 Each csv file is saved with following columns:
 - pc_name
 - eye
 - repetition, 
 - intensity
 - measurement_time,
 - spectrum
 
 The last column, spectrum, contains a list of several values, which are the values of the spectrum measured by the spectrometer. The number of values in the spectrum column is the same for all rows in the csv file. 
 
## Aim of thi script
We need to re-format this csv file so that the values in the column spectrum can actually be used for further analysis. The idea is to create a new csv file where each row corresponds to a single value in the spectrum column. The new csv file will have the following columns:
- pc_name
- eye
- repetition, 
- intensity
- measurement_time,
- spectral irradiance
Where spectral irradiance is a single value extracted from spectrum, corresponding to a specific wavelength 

```{r}
## Ensure the directory is set correctly first!

library(tidyverse)
# Load the CSV file 
cali_df_raw <- read_csv("RGMS-VR02_left_20250205T162306.csv")

# Process the spectrum column
cali_df_clean <- cali_df_raw %>%
  mutate(spectrum = str_remove_all(spectrum, "\\[|\\]")) %>%  # Remove brackets
  separate_rows(spectrum, sep = " ") %>%  # Expand into multiple rows
  mutate(spectrum = as.numeric(spectrum)) %>%  # Convert to numeric
  rename(spectral_irradiance = spectrum) %>% # Rename for clarity
  select(pc_name, eye, repetition, intensity, measurement_time, spectral_irradiance)  # Reorder columns and eliminate first one

# We do not have a column for wavelength, which we need to use the spectran software
cali_df_clean <- cali_df_clean %>%
  group_by(intensity) %>%
  mutate(wavelength = seq(from = 380, to = 780)) %>%  # Create a wavelength column from 380 to 780 at 1nm steps
  ungroup()

```
## Save outputs in different csv according to intensity
```{r}
# Get unique intensity values
intensities <- base::unique(cali_df_clean$intensity)

# Output folder
output_folder <- "H:/shine/code/experimental/calibration/results/processed/"

# Save separate CSV files for each intensity level
for (int in intensities) {
  df_filtered <- cali_df_clean %>% filter(intensity == int)  # Filter by intensity
  file_name <- paste0(output_folder, "intensity_", int, ".csv")  # Create filename
  # Save file
  write_csv(df_filtered,
            file_name)  
}

```

